---
title: "ExploratoryAnalysis"
format: pdf
editor: visual
---

## Q1: What does the data contain?

```{r}
library(ggplot2)
data <- read.csv("large_dataset.csv")
head(data)
sdata <- data[-c(1, 5, 8:25, 41:53)]
head(sdata)
str(data)
```

The data is from the (UFC Complete Dataset on Kaggle)\[<https://www.kaggle.com/datasets/maksbasher/ufc-complete-dataset-all-events-1996-2024>\] that is parsed from the (UFCStats website)\[<http://www.ufcstats.com/statistics/events/completed>\]. It has detailed information on all UFC fights from 1996-2024, and contains all individual fighter statistics as well as the in fight statistics. The dataset contains 95 variables. For the purposes of this project, we will be primarily focusing on the fighter statistics as predictor variables for fight outcomes. Fighter statistics include physical statistics such as height, weight, and stance, and career statistics such as strikes accuracy and takedown accuracy.

## Q2: What is the data variability like?

```{r}

ggplot(data = data, mapping = aes(x=method)) + geom_bar()

ggplot(data = data, mapping = aes(x=r_stance)) + geom_bar()


#difference in significant strikes between fighters by gender
ggplot(data = data, mapping = aes(x=sig_str_acc_diff, colour = gender)) + geom_freqpoly()
#key to seeing that the red/blue player won't influence results 

# red total average striking 
ggplot(data = data, mapping = aes(x=r_sig_str_acc_total, colour = gender)) + geom_freqpoly()

#red submission average
ggplot(data = data, mapping = aes(x=r_sub_avg, colour = gender)) + geom_freqpoly()

#red takedown average
ggplot(data = data, mapping = aes(x=r_td_avg, colour = gender)) + geom_freqpoly()


```

## Q3: am I missing any data?

```{r}
summary(sdata)
```

Most of the missing values appear in the difference columns, which we don't really have to use anyways.

## Q4: what type of covariation occurs within my variables?

```{r}
ggplot(data = data, mapping = aes(x=r_td_avg, y = r_sub_avg)) +geom_point()

ggplot(data = data, mapping = aes(x=r_sig_str_acc_total, y = r_sub_avg, colour = gender)) +geom_point()

ggplot(data = data, mapping = aes(x=r_sig_str_acc_total, y = r_losses_total, colour = gender)) +geom_point()

ggplot(data = data, mapping = aes(x=r_sig_str_acc_total, y = r_losses_total, colour = winner)) +geom_point()

ggplot(data = data, mapping = aes(x=r_sig_str_acc_total, y = winner)) + geom_boxplot()

ggplot(data = data, mapping = aes(x=r_sub_avg, y = winner)) + geom_boxplot()

ggplot(data = data, mapping = aes(x=r_sub_avg, y = r_stance)) + geom_boxplot()

```
